SMT-Boosted Security Types for Low-Level MPC

### 2024-10-10 esop25/intro.tex
checkpoint

Secure Multi-Party Computation (MPC) protocols support data privacy in
important modern, distributed applications such as privacy-preserving
machine learning
\cite{li2021privacy,knott2021crypten,koch2020privacy,liu2020privacy}
and Zero-Knowledge proofs in blockchains
\cite{ishai2009zero,lu2019honeybadgermpc,gao2022symmeproof,tomaz2020preserving}.
The security semantics of MPC include both confidentiality and
integrity properties incorporated into models such as real/ideal (aka
simulator) security and universal composability (UC), developed
primarily by the cryptography community \cite{evans2018pragmatic}.
Related proof methods are well-studied \cite{Lindell2017} but mostly
manual. Somewhat independently, a significant body of work in
programming languages has focused on definition and enforcement of
confidentiality and integrity \emph{hyperproperties}
\cite{10.5555/1891823.1891830} such as noninterference and gradual
release \cite{4223226,sabelfeld2009declassification}. Following a
tradition of connecting cryptographic and PL-based security models
\cite{10.1007/3-540-44929-9_1,10.1145/3571740}, recent work has also
recognized connections between MPC security models and hyperproperties
of, e.g., noninterference \cite{8429300}, and even leveraged these
connections to enforce MPC security through mechanisms such as
security types \cite{10.1145/3453483.3454074}. Major benefits of this
connection in an MPC setting include better language abstractions for
defining protocols and for mechanization and even automation of
security proofs.  The goal of this paper is to develop a PL model for
defining a variety of low-level probabilistic MPC protocols, to
formulate a collection of confidentiality and integrity
hyperproperties for our model with familiar information flow analogs,
and to show how these properties can be leveraged for improved proof
automation.

@@

The connection between information flow hyperproperties and MPC
security is also complicated especially at a low level.  MPC protocols
involve communication between a group of distributed clients called a
\emph{federation} that collaboratively compute and publish the result
of some known \emph{ideal functionality} $\idealf$, maintaining
confidentiality of inputs to $\idealf$ without the use of a trusted
third party. However, since the outputs of $\idealf$ are public, some
information about inputs is inevitably leaked. Thus, the ideal
functionality establishes a declassification policy
\cite{sabelfeld2009declassification}, which is more difficult to
enforce than pure noninterference.  And subtleties of, e.g.,
semi-homomorphic encryption are central to both confidentiality and
integrity properties of protocols and similarly difficult to track
with coarse-grained security types alone.

Nevertheless, as previous authors have observed
\cite{5a51987acaa84c43bb4bf5bcc7d01683}, low-level protocol design
patterns such as secret sharing and circuit gate structure have
compositional properties that can be independently verified and then
leveraged in larger proof contexts. We contribute to this line of work
by developing an automated verification technique for subprotocols and
show how it can be integrated as a tactic in a larger security proof.

@@

In summary, our work provides a complete methodology for end-to-end verification
of MPC protocols via three components:
%
\begin{enumerate}
\item A \textbf{low-level language} for defining MPC primitives
  (Section \ref{section-lang}) with an associated \textbf{metalanguage}
  to ease programming (Section \ref{section-metalang})\footnote{By metalanguage
  we mean a multi-stage aka metaprogramming language where code is a value, as
  in, e.g., MetaML \cite{TAHA2000211}.}. 
\item A \textbf{fully-automated verification method} for low-level MPC
  primitives in $\mathbb{F}_2$ (Section \ref{section-bruteforce}).
\item A \textbf{partially-automated verification method} for
  MPC protocols, which leverages automated proofs for
  low-level primitives (Section \ref{section-example-gmw}).
\end{enumerate}
%
As part of this methodology, we also develop hyperproperties that
encode MPC security that may be of independent interest. The complete
methodology enables the verification of real-world MPC protocols like
GMW~\cite{goldreich2019play}.

\paragraph{Language design.} In Section \ref{section-lang} we
develop a new probabilistic programming language $\minifed$ for
defining synchronous distributed protocols over an arbitrary
arithmetic field. The syntax and semantics provides a succinct account
of synchronous messaging between protocol \emph{clients}. In Section
\ref{section-metalang} we define a metalanguage $\metaprot$ that
dynamically generates $\minifed$ protocols. It is able to express
important low-level abstractions, as we illustrate via implementations
of protocols including Shamir addition (Section \ref{section-lang}),
GMW boolean circuits (Section \ref{section-example-gmw}), and Beaver
Triple multiplication gates with BDOZ authentication (Section
\ref{section-example-bdoz}).

\testcompfig

\paragraph{Hyperproperty formulation.} In Section \ref{section-model} we
develop our formalism for expressing the joint probability mass function of
program variables, and give standard definitions of passive and
malicious real/ideal security in our model. In Section
\ref{section-hyper}, we formulate a variety of familiar information
flow properties in our probabilistic setting, including conditional
noninterference, gradual release, and robust declassification, and
consider the relation between these and real/ideal security.  While it
has been previously shown that probabilistic conditional
noninterference is sound for passive security, we formulate new
properties of integrity which, paired with passive security, imply
malicious security (Theorem \ref{theorem-integrity}). We observe
in Section \ref{section-example-bdoz} that authentication mechanisms
such as BDOZ/SPDZ style MACs enforce a strictly weaker property
of ``cheating detection'' (Lemma \ref{lemma-cheating}).

\paragraph{Fully and partially automated verification.} In Section
\ref{section-bruteforce} we develop a method for automatically
computing the probability mass function (pmf) of $\minifed$ protocols
in $\mathbb{F}_2$, that can be automatically queried to enforce
hyperproperties of security. This method is perfectly accurate but has
high complexity; we show this can be partially mitigated by conversion
of protocols in $\mathbb{F}_2$ to stratified Datalog which is amenable
to HPC acceleration. Furthermore, in Section \ref{section-example-gmw}
we consider in detail how this automated technique can be used as a
local automated tactic for proving security in arbitrarily large GMW
circuits using conditional probabilistic independence as in
\cite{li2023lilac} (Lemmas \ref{lemma-gmwtactic} and
\ref{lemma-gmwinvariant} and Theorem \ref{theorem-gmw}).

@@

Our main focus is on PL design and automated and semi-automated
reasoning about security properties of low-level MPC protocols. Prior
work has considered \textbf{automated} verification of
\textbf{high-level} protocols and \textbf{manual} verification of
\textbf{low-level} protocols---but none offers the combination of
automation and low-level support we consider.
%
We summarize this comparison in Figure
\ref{fig-comp}, with the caveat that works vary in the degree of
development in each dimension.

@@

>>>>

Secure Multi-Party Computation (MPC) protocols support data privacy in
important modern, distributed applications such as privacy-preserving
machine learning
\cite{li2021privacy,knott2021crypten,koch2020privacy,liu2020privacy}
and Zero-Knowledge proofs in blockchains
\cite{ishai2009zero,lu2019honeybadgermpc,gao2022symmeproof,tomaz2020preserving}.
The security semantics of MPC include both confidentiality and
integrity properties incorporated into models such as real/ideal (aka
simulator) security and universal composability (UC), developed
primarily by the cryptography community \cite{evans2018pragmatic}.
Related proof methods are well-studied \cite{Lindell2017} but mostly
manual. To support automated reasoning, recent work has reformulated
MPC security semantics as \emph{hyperproperties}
\cite{8429300,10.1145/3453483.3454074,skalka-near-ppdp24}.
Separately, recent research in the SMT community has developed
theories of finite fields \cite{SMFF} with clear relevance to
verification of cryptographic schemes that rely on field
arithmetic. The goal of this paper is to combine these two approaches
in an automated type system for verifying correctness and security
properties of MPC protocols.

@@

@@

\begin{enumerate}
\item A \textbf{low-level language} for defining MPC primitives
  (Section \ref{section-lang}) with an associated \textbf{metalanguage}
  to ease programming (Section \ref{section-metalang})\footnote{By metalanguage
  we mean a multi-stage aka metaprogramming language where code is a value, as
  in, e.g., MetaML \cite{TAHA2000211}.}. 
\item A \textbf{fully-automated verification method} for low-level MPC
  primitives in $\mathbb{F}_2$ (Section \ref{section-bruteforce}).
\item A \textbf{partially-automated verification method} for
  MPC protocols, which leverages automated proofs for
  low-level primitives (Section \ref{section-example-gmw}).
\end{enumerate}

@@

\cnote{This sections needs updating with, including with
  \cite{skalka-near-ppdp24,SMFF}}.
Our main focus is on PL design and automated and semi-automated
reasoning about security properties of low-level MPC protocols. Prior
work has considered \textbf{automated} verification of
\textbf{high-level} protocols and \textbf{manual} verification of
\textbf{low-level} protocols---but none offers the combination of
automation and low-level support we consider.
%
We summarize this comparison in Figure
\ref{fig-comp-wrap}, with the caveat that works vary in the degree of
development in each dimension.

@@

\compwrapfig
